Parametric analysis of ambisonic audio

### 2020-07-05 Chapters/SELD.tex
tbc

\begin{figure}[th!]
  \centering
  \centerline{\includegraphics[width=0.75\textwidth]{Figures/SELD/ARCH.png}}
  \caption{Architecture of the proposed methodology.}
  \label{fig:scheme}
\end{figure}

@@

Table~\ref{tab:results} summarizes the results of the experiments using the recommended data split: training with folds 3 to 6, validation with fold 2 and testing with fold 1. 
% This structure has been promoted by the Challenge organization as a fair way for method comparison. 
% Furthermore, the usage of an unseen split 
% Furthermore, it is expected to provide similar results to the evaluation set, given that part of the data remains unseen at training.
Results are reported for three different systems: the baseline and the two proposed methods \textit{PAPAFIL1} and \textit{PAPAFIL2}.
The results of their respective oracle results, \textit{PAPAFIL1-O} and \textit{PAPAFIL2-O}, are also provided.

Both proposed approaches outperform the baseline system in three out of the four evaluation metrics ($\text{ER}_{20}$, $\text{F}_{20}$ and $\text{LE}_{CD}$). 
Although the results obtained by both of them are similar, \textit{PAPAFIL2} obtains better classification scores ($\text{ER}_{20}$ and $\text{F}_{20}$), and \textit{PAPAFIL1} performs subtly better regarding localization error ($\text{LE}_{CD}$).
However, the localization recall results ($\text{LR}_{CD}$) are slightly worst than the baseline in both cases. This fact does not prevent the proposed methods to have a $\text{SELD}$ score better than the baseline: 0.41 (\textit{PAPAFIL1}) and 0.38 (\textit{PAPAFIL2}), against 0.47 (\textit{BASELINE}).

The results obtained by the oracle methods are within the expected ranges. \textit{PAPAFIL1-O} performs almost perfectly regarding $\text{LE}_{CD}$, but the classification errors influence the $\text{LR}_{CD}$ result.
In turn, \textit{PAPAFIL2-O} performs better than \textit{PAPAFIL1-O} regarding all metrics, excepting $\text{LE}_{CD}$; this improvement is specially noticeable in $\text{LR}_{CD}$, with a performance difference of about 15\%.
The good results obtained by \textit{PAPAFIL2-O} validate the proposed particle filtering approach, an leave space for improvements that might be given by a better understanding and fine tuning of the model.

The performance of the proposed methods deteriorates noticeably with overlapping sounds.
% Although the performance is good in general terms with both \textit{PAPAFIL} approaches, the results deteriorate noticeably with overlapping sounds.
A closer inspection reveals that, in many occasions, the TF bins passing the single-source test mostly belong to one out of two simultaneous sources. 
It is a known issue that performance of DirAC diffuseness is reduced when two sources are present \cite{epain2016spherical}; similar problems have been reported in \cite{adavanne2019localization}, where an instantaneous source number estimator is used in combination with the particle filter.
As in that case, the results suggest the need for more sophisticated source detection and counting methods.

\begin{table}[th!]
\begin{footnotesize}
\caption{Evaluation results on the development set.}
    \begin{center}
    \begin{tabular}{cccccc}
    \toprule
    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
    \midrule
    \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & \textbf{60.7} \% & 0.47 \\
    \textit{PAPAFIL1} & 0.60 & 49.8 \% & \textbf{13.4}$^{\circ}$ & 54.4 \% &0.41\\

    \textit{PAPAFIL2} & \textbf{0.57} & \textbf{54.0} \% &13.8$^{\circ}$ & 59.7 \% & \textbf{ 0.38} \\

    \midrule
    % this is oracle beam
    \textit{PAPAFIL1-O} & 0.37 & 67.0 \% & 2.0$^{\circ}$ & 68.6 \% & 0.26 \\
    \textit{PAPAFIL2-O} & 0.32 & 79.6 \% & 8.5$^{\circ}$ & 82.4\% & 0.19 \\

    % \midrule
    % \textit{PAPAFIL1} &  0.57 & 55.6 \% & 15.6$^{\circ}$ & 66.7 \% & 0.36 \\
    % % PAPAFIL_1 & 0.47  & 64.1 \% & 14.1$^{\circ}$ & 70.1 \% & 0.30 \\ ?????????
    % % this is oracle beam
    % \textit{PAPAFIL1-O} & 0.08 & 93.7 \% &  0.2$^{\circ}$ & 94.0 \% & 0.05 \\

    %  \textit{PAPAFIL2} & \textbf{ 0.44} & \textbf{68.0} \% & \textbf{13.3}$^{\circ}$ & \textbf{79.6} \% & \textbf{0.26} \\

    % \textit{PAPAFIL2-O} & 0.41 & 71.1 \% & 12.3$^{\circ}$ & 82.0\% &  0.24 \\
    \bottomrule
    \end{tabular}
    \end{center}
    \label{tab:results}
\end{footnotesize}
\end{table}

@@

To conclude the analysis, Fig.~\ref{fig:importance} shows the relative importance of the fifteen most relevant acoustic features for the \textit{PAPAFIL2} classifier model.
Event duration is clearly the feature with the highest importance, and effective duration (duration of the signal discarding silence) also appears in the eighth position.
This fact can help to explain the better performance of \textit{PAPAFIL2} over \textit{PAPAFIL1}: the temporal activities of the events in training and prediction are much more similar to each other in the former method, as a consequence of the training set generation approach. 
In order to provide a deeper insight on the temporal characteristics of the data under consideration, the durations of all sound events in the \textit{PAPAFIL2} training set are summarized in Figure~\ref{fig:eventdurations}.

@@

Furthermore, it is interesting to notice the high relevance of low-level features, and specifically several MFCC combinations (eight of the fifteen reported features) and  various extractors related to the spectral structure. The absence of pitch, harmonic and envelope features in the list represents a significant finding as well.

\todo{TODO: DISCUSS RESULTS OF EVALUATION SET!!}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Introduction}
%\label{sec:intro}
%
%Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual  event present in a sound field, the temporal activity, spatial location, and sound class to which it belongs. SELD is a current research topic which deals with microphone array processing and sound classification, with potential applications in the fields of signal enhancement, autonomous navigation, acoustic scene description or surveillance, among others.
%
%SELD arises from the combination of two different problems: Sound Event Detection (SED) and Direction of Arrival (DOA) estimation. The number of works in the literature which jointly address SED and DOA problems is relatively small. It is possible to classify them by the type of microphone arrays used: distributed \cite{grobler2017sound, butko2011two, chakraborty2014sound} or near-coincident \cite{hirvonen2015classification, lopatka2016detection, Adavanne2018_JSTSP}.
%As mentioned in \cite{Adavanne2018_JSTSP}, the usage of near-coincident circular/spherical arrays enables the representation of the sound field in the spatial domain, using the spherical harmonic decomposition, also known as Ambisonics \cite{gerzon1973periphony, daniel2000representation}. Such spatial representation allows a flexible, device-independent comparison between methods. Furthermore, the number of commercially available ambisonic microphones has increased in recent years due to their suitability for immersive multimedia applications.  
%Taking advantage of the compact spatial representation provided by the spherical harmonic decomposition, several methods for parametric analysis of the sound field in the ambisonic domain have been proposed  \cite{pulkki2006directional, berge2010high, Politis2018, pulkki2018parametric}.
%These methods ease sound field segmentation into direct and diffuse components, and further localization of the direct sounds.
%The advent of deep learning techniques for DOA estimation has also improved the results of traditional methods \cite{Adavanne2018_JSTSP}. However, none of the deep learning-based DOA estimation methods explicitly exploits the spatial parametric analysis. This situation is further extended to the SELD problem, with the exception of \cite{lopatka2016detection}, where DOAs are estimated from the \textit{active intensity vector}  \cite{pulkki2006directional}.
%
%The motivation for the proposed methodology is two-fold.
%First, we would like to check whether the usage of spatial parametric analysis in the ambisonic domain can improve the performance of SELD algorithms.
%Second, temporal information derived by the parametric analysis could be further exploited to estimate event onsets and offsets, thus lightening the event classifier complexity; such reduction might positively impact algorithm's performance.
%
%In what follows, we present the methodology and the architecture of the proposed system (Section~\ref{sec:method}). Then, we describe the design choices and the experimental setup (Section~\ref{sec:expe}), and discuss the results in the context of DCASE2019 Challenge - Task 3 (Section~\ref{sec:results}). A summary is presented in Section~\ref{sec:conclusion}. 
%%In order to support open access and reproducibility, all code is freely available at \cite{code}. \todo{code ref}.
%
%
%\section{Method}
%\label{sec:method}
%
%The proposed method presents a solution for the SELD problem splitting the task into four different problems: \textit{DOA  estimation}, \textit{association}, \textit{beamforming} and \textit{classification}, which will be described in the following subsections.
%The former three systems follow a heuristic approach---in what follows, they will be jointly referred to as the \textit{parametric front-end}. Conversely, the \textit{classification} system is data-driven, and will be referred to as the \textit{deep learning back-end}. The method architecture is depicted in Figure~\ref{fig:diagram_all}.
%
%\todo{redo figures in this section}
%
%\begin{figure}[h]
%	\centering
%    \includegraphics[width=\columnwidth]{Figures//SELD/dcase_challenge_tech_report_all2.pdf}
%    \caption{System architecture.}
%    \label{fig:diagram_all}
%
%\end{figure}
%
%\subsection{DOA estimation}
%\label{ssec:doa_estimation}
%
%\begin{figure*}[h]
%
%    \includegraphics[width=\textwidth]{Figures/SELD/DOA1.pdf}
%    \caption{DOA estimation architecture.}
%    \label{fig:doa}
%
%\end{figure*}
%
%
%
%The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency spatial audio analysis.
%
%Let us consider a \textit{N3D}-normalized first-order ambisonic signal $x_n^m(t)$. 
%Using the DirAC parametric model, it is possible to obtain instantaneous time-frequency estimates of the Direction of Arrival $\Omega(k,n)$ (Eq.~\ref{eq:doa}), energy density $E(k,n)$ (Eq.~\ref{eq:energydensity}) and the diffuseness $\Psi(k,n)$ (Eq.~\ref{eq:psidefinition}).\\
%
%
%
%%Let us consider a first order ambisonic signal $$
%%Let us consider a first order ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
%%\begin{equation}
%%    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
%%\end{equation}
%%From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
%%\begin{equation}
%%    \begin{aligned}
%%    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
%%    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
%%    \end{aligned}
%%\end{equation}
%%where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
%
%It is desirable to identify the TF regions of $\Omega(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
%
%The first mask is the \textit{energy density mask}, which is used as an activity detector. A gaussian adaptive thresholding algorithm is applied to $E(k,n)$. This procedure selects TF bins with local maximum energy density, as expected from the direct path of the sources.
%
%The \textit{diffuseness mask} selects the TF bins with high energy propagation. Bins with a low diffuseness value represent a TF region where one sound source is predominant. A fixed threshold value $\Psi_{max}$ is used for the masking.
%
%The third mask is the \textit{DOA variance mask}. It tries to select TF regions with small standard deviation with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.\\
%
%The three masks are then applied to the DOA estimation, obtaining the TF-filtered instanteneous DOAs $\bar{\Omega}(k,n)$. 
%Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
%The median filter is applied in a TF bin belonging to $\bar{\Omega}(k,n)$ only if the number of TF bins belonging to $\bar{\\Omega}(k,n)$ in its vicinity is greater than a given threshold $B_{min}$.
%The resulting filtered DOA estimation is referred to as $\breve{\Omega}(k,n)$.
%
%
%\subsection{Association}
%\label{ssec:association}
%
%\begin{figure*}[h]
%    \includegraphics[width=\textwidth]{Figures/SELD//ASS.pdf}
%    \caption{Association architecture.}
%    \label{fig:association}
%
%\end{figure*}
%
%The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\Omega}(k,n)$ to a set of events, each one having a specific onset, offset and location. \\
%
%First, DOA estimates are resampled into \textit{frames} with the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
%An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\Omega}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is higher than a given threshold $K_{min}$.\\
%
%Next, the standard deviation in azimuth $\sigma_{\varphi}$ and elevation $\sigma_{\theta}$ of the frame-based DOA estimates $\breve{\Omega}(k,m)$ are compared to a threshold value $\sigma_{max}$, and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
%
%\begin{equation}
%\begin{aligned}
%	o(m) =  \begin{cases}
%    		1, &\text{if } \sigma_{\varphi}/2 + \sigma_{\theta} < \sigma_{max},\\
%    		2, &\text{otherwise}.
%    \end{cases}    		
%\end{aligned}
%\end{equation}
%
%The clustered values $\Omega_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\Omega}(k,m)$, using a modified version of K-Means which minimizes the central angle distance. Notice that, for $o(m)=1$, the operation is equivalent to the median.\\
%
%The following step is the grouping of clustered DOA values into events.
%Let us define $\Omega_S(m)$ as the frame-wise DOA estimations belonging to the event $S$. A given clustered DOA estimation $\Omega_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
%\begin{itemize}
%    \item The central angle between  $\Omega_{\text{cluster}}(m)$ and the median of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
%    \item The frame distance between $m$ and the closest frame of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
%\end{itemize}
%
%The resulting DOAs $\Omega_S(m)$ are subject to a postprocessing step with the purpose of 
%delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length. 
%
%Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\Lambda_S = (\Omega_S, \text{onset}_S, \text{offset}_S)$.
%
%\subsection{Beamforming}
%\label{ssec:Beamforming}
%
%The last step performed in the front-end is the input signal segmentation. The spatial and temporal information provided by the annotations $\Lambda_S$ are used to produce monophonic signal estimations of the events, $\tilde{x}_S(t)$, as the signals captured by a virtual first-order cardioid (Eq.~\ref{eq:decodingequation}):
%\begin{equation}
%\tilde{x}_S(t) = x_n^m(t) Y_n^m(\Omega_S) \alpha_n,
%\end{equation}
%with $\alpha_n = [1, 1, 1, 1]$ using the \textit{basic} decoding weights.
%
%
%
%\subsection{Deep learning classification back-end}
%\label{ssec:backend_method}
%
%%intro - problem formulation==============
%The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{s}_S(t)$.
%Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
%Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
%It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\Lambda_S$ are perfectly estimated.
%
%%2 stages: context
%The classification method is divided into two stages.
%First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
%Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $c \in \left\lbrace 1 ...C \right\rbrace$, with $C=11$.
%Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
%Only one label is predicted for each incoming event, such that there is no binarization stage needed as in a standard SED task.
%
%%archi=============
%The proposed CRNN is depicted in Figure \ref{fig:archi}.
%It presents three convolutional \textit{blocks} to extract local features from the input representation.
%Each convolutional block consists of one convolutional layer, after which the resulting feature maps are passed through a ReLU non-linearity \cite{nair2010rectified}.
%This is followed by a max-pooling operation to downsample the feature maps and add invariance along the frequency dimension.
%The target classes vary to a large extent in terms of their temporal dynamics, with some of them being rather impulsive (e.g., \textit{Door slam}), while others being more stationary (e.g., \textit{Phone ringing}). 
%Therefore, after stacking the feature maps resulting from the convolutional blocks, this representation is fed into one bidirectional recurrent layer in order to model discriminative temporal structures.
%Specifically, 64 nodes of gated recurrent units (GRU) are used with \textit{tanh} activations.
%The recurrent layer is followed by a Fully Connected (FC) layer, and finally a 11-way softmax classifier layer produces the event-level probabilities.
%Dropout is applied extensively.
%The loss function used is categorical cross-entropy.
%The model has $\sim$175k weights.
%
%\begin{figure}[ht]
%  \vspace{-3mm}
%  \centering
%  \centerline{\includegraphics[width=0.9\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
%  \vspace{-2mm}
%  \caption{Back-end architecture.}
%  \label{fig:archi}
%    \vspace{-5mm}
%\end{figure}
%
%
%
%
%%===============================================================================
%
%\section{Experiments}
%\label{sec:expe}
%
%\subsection{Dataset, evaluation metrics and baseline system}
%\label{ssec:dataset}
%We use the TAU Spatial Sound Events 2019 - Ambisonic, which provides first-order ambisonic recordings.
%Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
%The dataset features a vocabulary of $C=11$ classes encompassing human sounds and sound events typically found in indoor office environments.
%The dataset is split into a development and evaluation sets. 
%The development set consists of a four fold cross-validation setup.\\
%
%The SELD task is evaluated with individual metrics for the SED and DOA problems:
%\begin{itemize}
%	\item SED: F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments.
%	\item DOA: DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise.
%\end{itemize}
%The \textit{SELD score} is an averaged summary of the system performance.
%A more detailed description of the evaluation metrics can be found in \cite{Adavanne2018_JSTSP}.\\
%
%The baseline system features a CRNN that jointly performs DOA and SED through multi-task learning \cite{Adavanne2018_JSTSP}.
%Baseline results are shown in Table \ref{tab:results_real}.
%
%
%\subsection{Parametric front-end}
%\label{ssec:param_frontend}
%
%Based on the method's exploratory analysis, we propose the following set of parameter values, which are shown in Table~\ref{tab:params_front}.
%
%
%\begin{table}[!htbp]
%
%\centering
%\caption{Parameter values for the selected configuration. Top: \textit{DOA analysis} parameters. Bottom: \textit{Association} parameters.}
%\begin{tabular}{cccc}
%\toprule
%Parameter & Unit & Value\\
%\midrule
%sampling rate & Hz  & 48000\\
%STFT window size & sample  & 256\\
%STFT window overlap & sample & 128\\
%STFT window type & -  &  Hann\\
%analysis frequency range & Hz & [0,8000]\\
%time average vicinity radius $r$ & bin& 10 \\
%diffuseness mask threshold $\Psi_{max}$ & - & 0.5 \\
%energy density filter length & bin & 11\\
%std mask vicinity radius & bin &  2\\
%std mask normalized threshold & - & 0.15\\
%median filter minimum ratio $B_{min}$ & - & 0.5 \\
%median filter vicinity radius (k,n) & bin & (20, 20) \\
%\midrule
%frame size $h$ & s  & 0.02\\

>>>>

\begin{figure}[th!]
  \centering
  \centerline{\includegraphics[width=0.55\textwidth]{Figures/SELD/ARCH.png}}
  \caption{Architecture of the proposed methodology.}
  \label{fig:scheme}
\end{figure}

@@

\begin{table}[t]
\begin{footnotesize}
\caption{System evaluation. Top: results on the cross-validation development set. Bottom: results on the evaluation set.}
    \begin{center}
    \begin{tabular}{cccccc}
    \toprule
    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
    \midrule
    \textit{BASELINE} & 0.70   & 39.5 \% & 23.2$^{\circ}$ & \textbf{62.1} \% & 0.45 \\
    \textit{PAPAFIL1} & 0.60 & 49.8 \% & \textbf{13.4}$^{\circ}$ & 54.4 \% &0.41\\
    \textit{PAPAFIL2} & \textbf{0.57} & \textbf{54.0} \% &13.8$^{\circ}$ & 59.7 \% & \textbf{0.38} \\

    % this is oracle beam
    \textit{PAPAFIL1-O} & 0.37 & 67.0 \% & 2.0$^{\circ}$ & 68.6 \% & 0.26 \\
    \textit{PAPAFIL2-O} & 0.32 & 79.6 \% & 8.5$^{\circ}$ & 82.4\% & 0.19 \\
	\midrule
	 \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & 60.7 \% & 0.47 \\
    \textit{PAPAFIL1} & 0.55 & 56 \% & 12.8$^{\circ}$ & 61.1 \% &0.36\\
    \textit{PAPAFIL2} & \textbf{0.51} & \textbf{60.1} \% &\textbf{12.4}$^{\circ}$ & \textbf{65.1} \% & \textbf{ 0.33} \\

    \bottomrule
    \end{tabular}
    \end{center}
    \label{tab:results}
\end{footnotesize}
\end{table}

Table~\ref{tab:results} summarizes the results of the experiments for development and evaluation datasets. Results on the development set have been computed using the provided cross-validation scheme: split 1 for testing, split 2 for validation, and splits 3 to 6 for training.

Results are reported for three different systems: the baseline and the two proposed methods \textit{PAPAFIL1} and \textit{PAPAFIL2}.
The results of their respective oracle results, \textit{PAPAFIL1-O} and \textit{PAPAFIL2-O}, are also provided for the development set.

Regarding the development set, both proposed approaches outperform the baseline system in three out of the four evaluation metrics ($\text{ER}_{20}$, $\text{F}_{20}$ and $\text{LE}_{CD}$). 
Although the results obtained by both of them are similar, \textit{PAPAFIL2} obtains better classification scores ($\text{ER}_{20}$ and $\text{F}_{20}$), and \textit{PAPAFIL1} performs subtly better regarding localization error ($\text{LE}_{CD}$).
However, the localization recall results ($\text{LR}_{CD}$) are slightly worst than the baseline in both cases. This fact does not prevent the proposed methods to have a $\text{SELD}$ score better than the baseline: 0.41 (\textit{PAPAFIL1}) and 0.38 (\textit{PAPAFIL2}), against 0.47 (\textit{BASELINE}).

The results obtained by the oracle methods are within the expected ranges. \textit{PAPAFIL1-O} performs almost perfectly regarding $\text{LE}_{CD}$, but the classification errors influence the $\text{LR}_{CD}$ result.
In turn, \textit{PAPAFIL2-O} performs better than \textit{PAPAFIL1-O} regarding all metrics, excepting $\text{LE}_{CD}$; this improvement is specially noticeable in $\text{LR}_{CD}$, with a performance difference of about 15\%.

The good results obtained by \textit{PAPAFIL2-O} validate the proposed particle filtering approach, and leave space for improvements that might be given by a better understanding and fine tuning of the model.

The overall tendency is maintained in the evaluation set results. Both proposed methods outperform again the baseline, improving the development set SELD score in five points; since the baseline slightly decreases the performance for the evaluation set, the score difference with respect to the proposed methods diverges significantly.
\textit{PAPAFIL2} is the method that clearly performs better on the evaluation set, with better results than \textit{PAPAFIL1} in all evaluation metrics. \\

The performance of the proposed methods deteriorates noticeably with overlapping sounds.
% Although the performance is good in general terms with both \textit{PAPAFIL} approaches, the results deteriorate noticeably with overlapping sounds.
A closer inspection to the development set results reveals that, in many occasions, the TF bins passing the single-source test mostly belong to one out of two simultaneous sources. 
It is a known issue that performance of DirAC diffuseness is reduced when two sources are present \cite{epain2016spherical}; similar problems have been reported in \cite{adavanne2019localization}, where an instantaneous source number estimator is used in combination with the particle filter.
As in that case, the results suggest the need for more sophisticated source detection and counting methods.

%\begin{table}[th!]
%\begin{footnotesize}
%\caption{Evaluation results on the cross-validation development set.}
%    \begin{center}
%    \begin{tabular}{cccccc}
%    \toprule
%    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
%    \midrule
%    \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & \textbf{60.7} \% & 0.47 \\
%    \textit{PAPAFIL1} & 0.55 & 56 \% & 12.8$^{\circ}$ & 61.1 \% &0.36\\
%    \textit{PAPAFIL2} & \textbf{0.51} & \textbf{60.1} \% &\textbf{12,4}$^{\circ}$ & 65.1 \% & \textbf{ 0.33} \\
%    \bottomrule
%    \end{tabular}
%    \end{center}
%    \label{tab:results}
%\end{footnotesize}
%\end{table}

@@

Fig.~\ref{fig:importance} shows the relative importance of the fifteen most relevant acoustic features for the \textit{PAPAFIL2} classifier model.
Event duration is clearly the feature with the highest importance, and effective duration (duration of the signal discarding silence) also appears in the eighth position.
This fact can help to explain the better performance of \textit{PAPAFIL2} over \textit{PAPAFIL1}: the temporal activities of the events in training and prediction are much more similar to each other in the former method, as a consequence of the training set generation approach. 
In order to provide a deeper insight on the temporal characteristics of the data under consideration, the durations of all sound events in the \textit{PAPAFIL2} training set are summarized in Figure~\ref{fig:eventdurations}.

@@

Furthermore, it is interesting to notice the high relevance of low-level features, and specifically several MFCC combinations (eight of the fifteen reported features) and  various extractors related to the spectral structure. The absence of pitch, harmonic and envelope features in the list represents a significant finding as well. \\

The results of the DCASE 2020 Challenge for the SELD task are shown in Table~\ref{tab:dcase2020results}. The method entries corresponds to the best scoring method for each team submission, respecting the naming conventions established for the Challenge. Accordingly, \textit{PerezLopez} refers there to the \textit{PAPAFIL2} method. 
The final rank is computed as the cumulative rank across all evaluation metrics, sorted by ascending order. The data has been taken from the Task results website \cite{dcasetask}, which provides complete information regarding the challenge.

\begin{table}[t]
\begin{footnotesize}
\caption{DCASE 2020 Challenge Task 3 evaluation results}
    \begin{center}

\begin{tabular}{llccccc}
    \toprule
Rank       & Method  & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
    \midrule
\textbf{1} & Du & \textbf{0.20}  & \textbf{84.9} \%    & \textbf{6.0}$^{\circ}$   & 88.5 \%    &  \textbf{0.12 }    \\
2          & Nguyen          & 0.23  & 82.0 \%    & 9.3$^{\circ}$   &\textbf{ 90.0} \%    &  0.14     \\
3          & Shimada         & 0.25  & 83.2 \%    & 7.0$^{\circ}$   & 86.2 \%    &  0.15     \\
4          & Cao             & 0.36  & 71.2 \%    & 13.3$^{\circ}$  & 81.1 \%    &  0.22     \\
5          & Park            & 0.43  & 65.2 \%    & 16.8$^{\circ}$  & 81.9 \%    &  0.26     \\
6          & Phan            & 0.49  & 61.7 \%    & 15.2$^{\circ}$  & 72.4 \%    &  0.30     \\
7          & PerezLopez      & 0.51  & 60.1 \%    & 12.4$^{\circ}$  & 65.1 \%    &  0.33     \\
8          & Sampathkumar    & 0.53  & 56.6 \%    & 14.8$^{\circ}$  & 66.5 \%    &  0.35     \\
9          & Patel           & 0.55  & 55.5 \%    & 14.4$^{\circ}$  & 65.5 \%    &  0.38     \\
10         & Ronchini        & 0.58  & 50.8 \%    & 16.9$^{\circ}$  & 65.5 \%    &  0.39     \\
11         & Naranjo-Alcazar & 0.61  & 49.1 \%    & 19.5$^{\circ}$  & 67.1 \%    &  0.38     \\
12         & Song            & 0.57  & 50.4 \%    & 20.0$^{\circ}$  & 64.3 \%    &  0.38     \\
13         & Tian            & 0.64  & 47.6 \%    & 24.5$^{\circ}$  & 67.5 \%    &  0.40     \\
14         & Singla          & 0.88  & 18.0 \%    & 53.4$^{\circ}$  & 66.2 \%    &  0.58     \\
15         & Baseline        & 0.69  & 41.3 \%    & 23.1$^{\circ}$  & 62.4 \%    &  0.45     \\
\bottomrule      
\end{tabular}
\end{center}
    \label{tab:dcase2020results}
    \end{footnotesize}
\end{table}

It is important to remark that, apart from our proposed method, all other systems rely completely on deep learning algorithms, specially making use of different configurations and combinations of CRNNs. 
The only exception to that tendency is the method presented by Nguyen and colleagues, which is an adaptation of their recently proposed methodology \cite{Nguyen2020_task3_report, nguyen2020sequence}. In this method, localization is also based on the identification of the single-source TF bins; instead of DirAC diffuseness, they use a subspace analysis to find the TF bins associated with nearly rank-1 spatial covariance matrices (SCM). 
Furthermore, the instantaneous DOA estimates are obtained by finding the azimuth and elevation histogram peaks, with the steering directions obtained from the covariance matrix diagonals. 
While this method might contribute to provide a smaller localization error compared with our approach (9.3$^{\circ}$ versus 14.2$^{\circ}$ in $\text{LE}_{CD}$), it is more expense computationally, since it involves building a SCM for each TF bin, plus peak picking in two 1D histograms (one for each spherical coordinate). 
However, the $\text{LE}_{CD}$ results are also influenced by the classification accuracy; a deeper analysis would be required to perform a direct comparison of both approaches.

Table~\ref{tab:dcase2020results} also highlights the good result obtained by our method regarding localization error: it is the fourth best method with respect to $\text{LE}_{CD}$. 
Moreover, the two model-based localization methods (our system and Nguyen) rank fourth and second in the $\text{LE}_{CD}$ classification, respectively. 
This result suggests that parametric spatial audio analysis, combined with a robust tracking system, is able to produce state-of-the-art localization results in the context of SELD. \\

To conclude the analysis, Figure~\ref{fig:importance} shows the relationship between the system complexity, measured in number of parameters of the model, against $\text{SELD}$ score, for all submitted systems, denoted by their ranking position. 
The overall tendency is clear: the higher the number of parameters, the better (smaller) the $\text{SELD}$ score; the regression curve, computed from all observations, confirms such tendency. 
It must be noticed that the presented method (\textit{7} in the Figure) lies outside the 95\% confidence interval of the regression curve (shaded area). In other words, its $\text{SELD}$ score is significantly better than the tendency observed from all submissions, considering the system complexity. 
Figure~\ref{fig:importance} also highlights the great difference in complexity among submissions. Between our method (the less complex, with 20k parameters) and the first scoring method (the most complex, with 123M parameters) there are four magnitude orders of different, constituting a remarkable difference. The logarithmic mean of all submitted systems' complexities is around 2.3M parameters.

\begin{figure}[t]
  \centering
  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/seld_complexity_v4.png}}
  \caption{DCASE 2020 Task 3 submissions: complexity versus SELD score.}
  \label{fig:importance}
\end{figure}

@@
